Visualizing Project Evolution Through Abstract Syntax Tree Analysis

### 2016-04-10 paper/paper.tex
Updated tex.

% As a general rule, do not put math, special symbols or citations
% in the abstract
\begin{abstract}
What is a developer's contribution to a repository? By using only counting commits and number of lines changed, existing tools that analyze source code repositories fall short on showing the effective contributions made by each developer. When many commits are viewed as a group, the details are lost. Commit information can be misleading since lines of code give no indication of what was actually being worked on without careful examination of the changed code. Providing a semantic view of this information could provide deeper insights into how software projects evolve since changes to design and features are not clearly visible from line changes alone. We present TypeV: a method for visualizing Java source code repositories. Instead of counting line changes in a commit we extract detailed type information over time by using the differences between abstract syntax trees (ASTs). We are then able to track the additions and deletions of declarations and invocations for each type. Furthermore, we can track each author’s type usage over time. Using this tool we examine specific cases in well-known repositories where our tool reveals interesting and useful information. We then compare type coverage information from the AST compared to file coverage to determine if unique information is provided by type information.
\end{abstract}

@@

\begin{figure*}[!t]
\centering
\includegraphics[width=\textwidth]{antlr-3-code-freq-edit}
\caption{}
\label{fig:GH_code_freq}
\end{figure*}

@@

@@

In the following sections of the paper we will introduce a tool, TypeV, that we developed that visualizes changes to types used in a project over time. TypeV is able to extract type information from a repository by computing commit-by-commit abstract syntax tree differencing, which we describe in section~\ref{sec:methodology}. We seek to answer the following questions: \\

\textbf{RQ1:} How can we extract type and method usage over time from repositories? \\ \\
\textbf{RQ2:} Does the type information give any interesting or useful insight into the project? \\ \\
\textbf{RQ3:} How does type coverage compare to file coverage?

@@

\begin{adjustwidth}{2.5em}{0pt}
\#DECLARE $\vert$ INSERT $\vert$ java.net.URI $\vert$ 2 \\
\#DECLARE $\vert$ DELETE $\vert$ java.lang.String $\vert$ 2 \\
\end{adjustwidth}

@@

To allow for a more detailed analyzes of a specific component of the repository we added some basic filtering tools. The first is that a user can filter by type name. The second allows the user to view changes that fall in a specific start and end date. The third allows them to specify the time period of each bar and this includes hour, day, week, and month. Finally, they can choose to see only declarations, types used in declarations and invocations, or invocations.

>>>>

% As a general rule, do not put math, special symbols or citations
% in the abstract
\begin{abstract}
What is a developer's contribution to a repository? By only counting commits and number of lines changed, existing tools that analyze source code repositories fall short on showing the effective contributions made by each developer. When many commits are viewed as a group, the details are lost. Commit information can be misleading since lines of code give no indication of what was actually being worked on without careful examination of the changed code. Providing a semantic view of this information could provide deeper insights into how software projects evolve since changes to design and features are not clearly visible from line changes alone. We present TypeV: a method for visualizing Java source code repositories. Instead of counting line changes in a commit we extract detailed type information over time by using the differences between abstract syntax trees (ASTs). We are then able to track the additions and deletions of declarations and invocations for each type. Furthermore, we can track each author’s type usage over time. Using this tool we examine specific cases in well-known repositories where our tool reveals interesting and useful information. We then compare type coverage information from the AST compared to file coverage to determine if unique information is provided by type information. 
\end{abstract}

@@

\begin{figure*}[!t]
\centering
\includegraphics[width=\textwidth]{antlr-3-code-freq-edit}
\caption{GitHub graph showing code frequency. Note that the graph was truncated to better show the details. Thus the changes that occur in mid 2015 are much larger than appear. The additions reach over 400k and the deletions go over 300k.}
\label{fig:GH_code_freq}
\end{figure*}

@@

Consider tools easily available to an open source software developer: GitHub graphs~\cite{github-graphs}. Since GitHub tracks contributions based solely on number of commits, and numbers of lines changes, data may be lacking, or even misleading. In the case of the Antlr4 project (Figure~\ref{fig:GH_code_freq}), the graph creates more questions than it answers. What happened at those big spikes of line changes in the graph? One of the spikes is so large, we had to truncate the graph to be able to show it. How did these large changes affect the code base? Who committed these large changes? Was it a team of contributors or a one particularly prolific programmer? Plotting the number of lines changed may be misleading, because a huge spike in additions may be caused a number of reasons, such as the inclusion of automatically-generated documentation, the inclusion of an external library, or perhaps even a huge refactoring effort.

@@

In the following sections of the paper we will introduce a tool, TypeV, that we developed that visualizes changes to types used in a project over time. TypeV is able to extract type information from a repository by computing commit-by-commit abstract syntax tree differencing, which we describe in section~\ref{sec:methodology}. We seek to answer the following questions: \\ \\
\textbf{RQ1:} How can we extract type and method usage over time from repositories \\ \\
\textbf{RQ2:} Does the type information give any interesting or useful insight into the project? \\ \\
\textbf{RQ3:} How does type coverage compare to file coverage?

@@

\begin{verbatim}
#DECLARE | INSERT | java.net.URI | 2
#DECLARE | DELETE | java.lang.String | 2
\end{verbatim}

@@

To allow for a more detailed analysis of a specific component of the repository we added some basic filtering tools. The first is that a user can filter by type name. The second allows the user to view changes that fall within a specific start and end date. The third allows them to specify the time span represented by each bar: either hour, day, week, or month. Finally, they can choose to see only types declarations, types used in declarations and invocations, or only invocations.
